Closing the gap between statistical and scientific workflows for improved forecasts in ecology

### 2025-07-09 revision/response_r1.tex

\documentclass[11pt,letter]{article}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\usepackage{graphicx} % Required for inserting images
\usepackage{xcolor} 
\definecolor{darkgrey}{HTML}{4d4d4d}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{gensymb}
\usepackage{xr-hyper}
\makeatletter
  \long\def\myempty{}
  \def\XR@addURL#1{\XR@@dURL#1\myempty{}{}{}{}{}\\}
  \def\XR@@dURL#1#2#3#4#5#6#7\\{%
    {#1}{#2}%
    \ifx\myempty#6\@empty
      {#3}{#4}{\XR@URL}%
    \else
    \fi
}
\makeatother
\usepackage{hyperref}
\hypersetup{hidelinks,breaklinks=true}
\usepackage{cleveref}
\usepackage[labelfont=bf]{caption}
\bibliographystyle{../amnat}

\RequirePackage[labelfont={bf,sf},%
                font={small, sf}]{caption}

\RequirePackage[labelfont={bf,sf},%
                font={small, sf}]{caption}

\externaldocument{forecastflows_r1}
\newcommand{\lref}[1]{}
% (\href{file:forecastflows_r1\#lintarget:SOMELABEL}{line \ref*{lin:SOMELABEL}})

\DeclareEmphSequence{\itshape\color{darkgrey}}
\usepackage[most]{tcolorbox}
% \newtcolorbox{mybox}[1][]{%
%     blanker, 
%     left skip=0em,
%     left=1em,
%     borderline west={1pt}{2pt}{darkgrey},
%     breakable,
%     #1}
\newtcolorbox{mybox}[1][]{%
    blanker, 
    left skip=0em,
    left=0em,
    borderline west={0pt}{0pt}{white},
    breakable,
    #1}

\begin{document}
\setlength{\parindent}{0cm}
\setlength{\parskip}{7pt}

Editor and reviewer comments (we provide below the full context of the two reviewers' comments) are in \emph{italics}, while our responses are in regular text. \\

{\bf Response to editor's comments:}

\begin{mybox}
\emph{...}
\end{mybox}

We are happy to hear that both reviewers recognized the value of our work and we appreciate the opportunity to revise our manuscript. We have worked to clarify the points raised by the editor and the two reviewers in this new version.

{\bf Response to reviewers' comments:}

{\bf Reviewer 1}

\begin{mybox}
\emph{This opinion manuscript (MS) argues that adapting a modern statistical workflow will help resolve some key issues in forecasting models used in ecology to inform policy about biodiversity loss and climatechanges impacts. Specifically, the authors identity two specific “disconnected paradigms” for making such forecasts and argue that a workflow adopted from the Bayesian literature, and that emphasizes data simulation, would help unify these paradigms and make for a better scientific understanding of these important issues. I found the MS well-written, concise and easy to read, and well-organized with clearly laid out challenges and understandable solutions. I agree with the central proposal that the adoption of a “principled” workflow would be beneficial in many areas of ecology, and these types of studies in particular. I have no experience with these specific types of ecological models, but I do have experience with the Bayesian workflow underlying their key recommendation. So, I will focus my review on the recommendation and not the challenge. Below I outline several major concerns/questions, and below that minor comments and corrections.}
\end{mybox}

Response.

\begin{mybox}
\emph{\textbf{Inadequate justification of simulation.}\\
Simulating data and fitting models is key to the step-by-step workflow but I think the MS lacks sufficient
detail and motivation. The methods described come from the Bayesian literature, where simulating
parameters from priors (prior predictive/pushforward) or posterior (posterior predictive or retrodictive)
and then data, are natural. But a quick peek at the cited literature suggests that Bayesian models are not
used. So how would one adapt them? On L76 they say “... by fixing parameters to some reasonable
range of values (which is straightforward if the parameters are interpretable)” but I disagree with this.
Do you suggest analysts develop priors (using, e.g., Betancourt’s cited workflow case study), and then
simulate parameters from those, but ignore them in the subsequent fitting to empirical data? Or take a
range for each parameter and simulate parameters across N equal steps for each one? Posterior
predictive checks (retrodiction) are known to be conservative due to the double-use of the data (see
Conn et al. 2018 [https://doi.org/10.1002/ecm.1314]). How does that affect your recommendations and
how the analyst should use this tool?}
\end{mybox}

Response.

\begin{mybox}
\emph{Furthermore, I think the authors should include a simple illustration of how these retrodiction checks can flag the need for model changes. For instance, I imagine they could take a fitted model from one of the key cited papers like the LPI trend, and simulate data given spatial/phylogenetic structure, but then refit a model without that and show how the model fails to fit the data and how the analysts would
identify (visually, with a new figure) and resolve that by adding model structure/complexity. I am a big proponent of prior/posterior predictive checking and do not disagree with their recommended usage, rather I think more detail is needed for the MS. I think a judicious, simple example here could really help the reader interpret the recommendations and start thinking about how to apply it in their studies. Given the importance of this step in the overall thesis of the MS I believe more detail on how an analyst would do this would be worthwhile.}
\end{mybox}

Response.

\begin{mybox}
\emph{\textbf{Resolving process-based model issues.}\\
I do not see how the proposed workflow would resolve what appears to a key issue in the process-based
research, namely that individual model components are done separately. Box (B) is supposed to address
this but I am left unconvinced how this would resolve that. As I understand it, their workflow would lead
to improved submodels, but would not change how those were linked together. Perhaps I missed a
point here (admittedly I am unfamiliar with these models), but if not, I encourage the authors to try to
make this point clearer. Otherwise will the workflow “bridge the gap” between the paradigms?}
\end{mybox}

Response.

\begin{mybox}
\emph{\textbf{Causal analysis as a framework to develop research questions and elicit expertise.}\\
The authors repeatedly highlight (rightfully so) that more time needs to be spent developing the research question. E.g., L266 “By focusing on model development more tightly tied to ecological expertise…”. However, they give no advice on how to do that. I pose that the causal analysis literature can and should be considered as a good source to bolster this argument and guide readers, as it provides a robust scientific framework for this step of the workflow. For instance, Grace and Irvine (2020) [https://doi.org/10.1002/ecy.2962] is an accessible introduction to this topic with ecological examples. I suggest the authors either incorporate this literature into their recommendation, or justify why not. As Grace and Irvine argue, it is not really fair to ask readers to develop these hypotheses without giving advice on how to do just that.}
\end{mybox}

Response.

\begin{mybox}
\emph{\textbf{Improvements to Figure 1:}\\
I like this figure but improving the caption would really help the reader take in the material more efficiently. I suggest moving the upper right panel to the upper left, as it serves as a sort of legend of the three other panels and should be read first. I would label this one A as well. What is the significance of the figure-8 in B? How is that different than a circle? This caption does not define or discuss the panels A,B,C, and I think that it should. Why are the terms “pre-data” and “pre-model” only used in the capture and nowhere else in the MS? “Retrodictive checks” may need to be further clarified in the caption.}
\end{mybox}

Response.

\begin{mybox}
\emph{“..where a model is built..”\\
“..some feedback to..”}
\end{mybox}

\begin{mybox}
\emph{\textbf{Minor}\\
Page 1, Line 40 in abstract: what is “unified” about the workflow?}
\end{mybox}

\begin{mybox}
\emph{Page 1, Line 45 in abstract: drop “for”}
\end{mybox}

\begin{mybox}
\emph{L6: “date, alongside”}
\end{mybox}

\begin{mybox}
\emph{L36: These workflow citations are all unpublished (arXiv). I’m fine with that, but consider citing some from other fields that have gone through peer review. The caption of Figure 2 includes more, suggest putting them here too.}
\end{mybox}

\begin{mybox}
\emph{L103: What does “change the model structure” mean here? Adding constraints and priors? Removing or adding complexity?}
\end{mybox}

\begin{mybox}
\emph{L110: delete “then”}
\end{mybox}

\begin{mybox}
\emph{Fig 2: Caption should mention this is expanded from Fig. 1C.}
\end{mybox}

\begin{mybox}
\emph{L116, L168: I disagree that “reduce uncertainties” is the goal – instead it is to more accurately quantify it. See L 257 “properly accounting for their uncertainty” for instance.}
\end{mybox}

\begin{mybox}
\emph{L143: FYI the fisheries literature has used the terms pre-data post-model etc. for decades now and have implemented this simulation approach. E.g., https://journal.iwc.int/index.php/jcrm/article/view/239/17. I don’t think it’s necessary to cite this but worth being aware of. You might want to re-cite Hilborn and Mangel (1997) here as they explicitly call for more simulation testing in ecology.}
\end{mybox}

\begin{mybox}
\emph{L201: “..workflow, non…” This sentence partially answers my question about L103 above}
\end{mybox}

\begin{mybox}
\emph{Box A: Please check the logic of the first sentence. More papers because of less time confronting their models? Is that what you meant to say? “For examples” should be “For example”?}
\end{mybox}

\begin{mybox}
\emph{Box B: Please clarify the last sentence if possible. I’m not sure what “inappropriate way to accommodate their complexity” means here.}
\end{mybox}

\begin{mybox}
\emph{L225: I like this paragraph and it makes a point that I hadn’t seen written out so clearly.}
\end{mybox}

\begin{mybox}
\emph{L245: I feel like section 4.2 is slightly out of place. I’d rather see more text devoted to retrodictive checks than this. Consider dropping it and focusing on just a conclusion here}
\end{mybox}

{\bf Reviewer 2}

\begin{mybox}
\emph{The paper is well written and in a style that is accessible to a wide audience. The topic fits
under the theme of “workflow for applied data analysis” by proposing a workflow that if used
thoughtfully could lead to less divergence in forecasts and therefore more trust in the scientific
process. The proposed workflow is similar to the Bayesian workflow outlined in Gelman et
al. (2020) and is translated to two different approaches to ecological forecasting. I think
the proposed workflow could be applied or modified to be used with a wide range of research
questions within ecological forecasting if careful attention is paid to nuances and idiosyncrasies
associated with each unique research question, and I think this necessary attention to question-
specific nuances could be highlighted more in the paper. There are places where I feel more
specificity (or clarity) with respect to these nuances could improve the manuscript and its
potential for impact with respect to building trust in the scientific process. There is also a
lack of specificity in how, exactly, this type of workflow could be implemented because there
is no example application to data. I have outlined these areas under “Major comments” and
provided minor editorial comments in “Minor Comments.”}
\end{mybox}

% Note: I am an applied statistician with experience collaborating with ecologists and developing
% statistical methodology for large-scale long-term monitoring programs. The bulk of my work
% is more concerned with explanatory than predictive inference and my knowledge of building
% ecological forecasting models for inference on a global scale is limited. I reviewed the paper
% through this lens.

\begin{mybox}
\emph{\textbf{Major comments}\\
1. Clarity on inferential goals of “disconnected paradigms”\\
 When the paper discusses “harmonizing both trend estimation and forecasting” what is meant by this? The
title of the paper suggests the inferential goals from both approaches to be predictive
in nature (i.e., “forecasting trends”), but it seems plausible that the simpler models are
developed for more of an explanatory purpose (e.g., investigating relationships between
covariates and mean response) rather than predicting or forecasting future trends, which
would make the questions being asked by each modeling approach fundamentally dif-
ferent (see, Shmueli 2010). The paper could be improved if more clarity was provided
on the inferential goals of both types of modeling paradigms it discusses, and perhaps
provide some discussion why they need to be (or if they should be) harmonized if they
are in pursuit of different types of inference.}
\end{mybox}

\begin{mybox}
\emph{2. Can this workflow be applied to non-Bayesian models\\
The emphasis on simulation-based investigation of model components is highlighted as a key feature of
the workflow, so it seems a generative (or at least partially generative) model is assumed. Does this model need to be Bayesian? Explicit connections to posterior predictive checks and more discussion for how both prior- and posterior-predictive assessment could be used in a non-Bayesian context would strengthen this work. (Also note that prior predictive checks should be made explicit too – see minor comments).}
\end{mybox}

\begin{mybox}
\emph{3. Discussion of Calibration:\\
Lines 98-104 seem to be referring to a check on estimating
“the model” using synthetic data that are generated from the model (either through
optimization or McMC), but there is no mention of various computational approaches
that may be taken to fit “the model.” How does the chosen computational strategy fit
into the workflow if at all? Further, is the proposed workflow to conduct a full-blown
simulation calibration study (i.e., simulate-fit-check for capture) many times and across
all possible parameter values? Assuming yes, I’d be interested in how something like
that would scale for the types of models that are the focus of the paper? Could the
authors provide more guidance on how to conduct such a study in this context?}
\end{mybox}

\begin{mybox}
\emph{4. Gaps in the workflow\\
A big argument in the paper is that the proposed workflow
could lessen discrepancies in results between studies of trends (lines 123-128). However,
in the workflow there is no mention of transparent communication of study design, which
if ignored in analysis, could result in discrepancies in results for studies with similar
research questions. Differences in sampling designs must also be considered prior to
integrating different data sources (if integration is even possible). The importance of
study design/sampling design and its ties to appropriate modeling is not discussed. Is it
implicitly assumed that there is alignment between design, question, and model in this
workflow? Some mention of the importance of study design would strengthen the work.}
\end{mybox}

\begin{mybox}
\emph{Additionally, the choice of prior and posterior predictive checks (PPC) is non-trivial and
depends on the context of the study. There is discussion of these types of simulation-
based investigations, but no mention of the importance of developing checks that assess
key features of the observed data that are tied to the research question. Acknowledging
the complexity of this step, and including some ideas for how to do this in the context
of the types of problems discussed in the paper could strengthen the paper.}
\end{mybox}

\begin{mybox}
\emph{5. Theoretical examples provided, but no data application of the proposed workflow\\
Sections 3.1 and 3.2 and the Box outline theoretical examples of how the workflow
could be applied for trend estimation and forecasting in the context of types of data, but
there are no concrete data examples. A comment on how the proposed workflow could
be applied to real data and how it would scale with complexity of data and research
questions would be a good addition to the paper if it is not feasible to include real-data
examples with the general discussion.}
\end{mybox}

\begin{mybox}
\emph{6. Adopting the workflow\\
ecommendations are made to train ecologists more quantitatively, but there is no suggestion or encouragement of collaboration with statisticians. These types of workflows are generally part of statistical practice, so could an alternative to more quantitative training (which would have to come at the cost of some other aspect of training) be more cross-disciplinary collaboration with statisticians?}
\end{mybox}

\begin{mybox}
\emph{\textbf{Minor comments}\\
Abstract, line 45: remove “for” in, “could transform for ecological modeling.”}
\end{mybox}

\begin{mybox}
\emph{Line 52: change “in term” to “in terms”}
\end{mybox}

\begin{mybox}
\emph{Figure 1 caption, line 34: change “where a model is build” to “where a model is built”}
\end{mybox}

\begin{mybox}
\emph{Lines 75-81, the paper discusses a form of prior predictive checking without explicitly mentioning the connection. The connection should be mentioned and perhaps}
\end{mybox}

\begin{mybox}
\emph{Line 80: typo in “workflow” in “(see example worklow)”}
\end{mybox}

\begin{mybox}
\emph{Line 86: change “also sometimes called a posterior predictive check” to “in a Bayesian workflow this is called a posterior predictive check.”}
\end{mybox}

\begin{mybox}
\emph{Line 87: could be more specific about “model output” as typically a PPC investigates how the entire posterior predictive distribution for the response or some summary statistic that is a function of the response and parameters compares to the observed data (or observed summary stat computed from observed data).}
\end{mybox}

\begin{mybox}
\emph{Line 117: should “hypothesis” be “hypotheses”?}
\end{mybox}

\begin{mybox}
\emph{Lines 198-201: if the goal is accurate and relatively-precise forecasts, do all of the nuisance
parameters need to be identifiable for a model to be useful? That is, does identifiability
of all parameters matter for process-based models with a goal of forecasting?}
\end{mybox}

\bibliography{forecastflows}

\end{document}

### 2025-07-10 revision/forecastflows_r1.tex
start working on reviewer response
I was hoping to have more time for this today

\documentclass[11pt]{article}
\usepackage[top=0.70in, bottom=.80in, left=1in, right=1in]{geometry}
\renewcommand{\baselinestretch}{1.1}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{multicol}
\usepackage{tcolorbox}
\usepackage{caption}
\usepackage{todonotes}
\usepackage[font=sf]{caption}
\usepackage{hyperref}

@@

\noindent \emph{For:} Scientific and Statistical Workflow theme issue for \emph{Phil Trans A} as an \emph{Opinion}
% They mention a tex template here: https://royalsocietypublishing.org/rsta/for-authors#question4
% Word length: I should check what they emailed but they say no more than 13 printed pages with 650 per page (whoa, we should be well below that! I am thinking 3-5K would be good)
% Already 5K!

@@

\begin{figure}[h]
	\centering
	\includegraphics{figures/modeldataspaces_rotation}
	\caption{Different trajectories through model-data space during model development. Ideally, this begins in the pre-model/pre-data quadrant, where the research question is defined. The next step moves to the post-model/pre-data quadrant, where a model is build---independently of the data. This is where we should spend a good deal of time evaluating model behavior using simulated data. Once we are satisfied with our model, we can move to the post-model/post-data quadrant, where the model is fitted to real data. Then, we perform retrodictive checks to compare model predictions to observations, which likely give us some feedbacks to refine our model.}
	\label{fig:modeldata}
\end{figure}

@@

We argue a workflow that moves along the data-model space in a coherent sequence of steps (figure \ref{fig:modeldata}c) could reduce many of these problems and thus improve ecological science.
The first step of this workflow is to define an explicit research question and formulate hypotheses (step 1, figure \ref{fig:workflow}). This involves making clear assumptions about the most influential drivers, within the specific context of our study. This step should guide the construction of a narrative model of how we believe the system works, focusing on the mechanisms that could generate the data we observe, including the observational error. 
From this narrative, we can then develop a mathematical model---an ensemble of equations that encapsulates our knowledge and is designed to answer our research question (step 2, figure \ref{fig:workflow}). The general idea is to start with a relatively simple model that we could refine later (see example worklow we provide). At this stage, prioritizing biologically meaningful parameters is crucial, as it allows us to have a sense of plausible parameter values. This means choosing a model formulation where each parameter corresponds to an interpretable behavior (which sometimes requires considering alternative parameterizations).

With a model in place, the next step focuses on testing and understanding it via data simulation (step 3, figure \ref{fig:workflow}). `Fake' or `test' data are generated directly from the model by fixing parameters to some reasonable range of values (which is straightforward if the parameters are interpretable) and from fake predictor data. 
We then fit our model to this simulated dataset and evaluate its ability to recover the prescribed values. At this stage, the focus is on understanding the model, so we may need to spend time thinking about whether each parameter is reasonable or not. This is also a step for making sure the model is working as expected (see example worklow).

Once we are confident about our model structure, we can introduce real data as part of an initial model fitting step (step 4, figure \ref{fig:workflow}). This way, we obtain parameter estimates constrained by observations. 
These parameter estimates lead to the second data simulation step, this time using our fitted model parameters to generate predictions (step 5, figure \ref{fig:workflow}). This---which we call a retrodictive check (\citealp{betanworkflow}, also sometimes called a posterior predictive check, \citealp{Gelman2020})---allows model output to be compared to observations. 
It's only once all steps have been completed that we can interpret parameter values with respect to our research question.
The workflow encourages a focus on the full model, where any parameter (such as a trend estimate) must be carefully interpreted alongside others, as all are fundamental components that shape both inference and forecasting.

@@

\begin{figure}[h]
	\centering
	\vspace*{-0.1cm}
    \hspace*{-1.5cm}
	\includegraphics{figures/figure_worflow_wsteps}
	\caption{The workflow we propose here \citep[which builds from recent advances in workflows,][]{betanworkflow,Gelman2020,Schad2020,grinsztajn2021,vandeschoot2021,Wolkovich2024} focuses on iterative feedbacks between the research question, model and data. Most research should start at Step 1, with the research question, followed by extensive model building and evaluation using simulated data (Steps 2 and 3) before proceeding to fitting the model on empirical data and examining it through retrodictive checks (Steps 4 and 5). Within this workflow, forecasting is a natural output of the process and not a separate process or one available for only certain modeling approaches.  The integration of observations (the data stream) occurs at Step 4---only after the model has been thoroughly evaluated---and can also highlight opportunities to collect new data and enrich the data stream.}
	\vspace*{-0.5cm}
	\label{fig:workflow}
\end{figure}

@@

\centerline{\bf Trend estimation}
\vspace*{2mm}
\begin{minipage}[t]{\linewidth}
    \includegraphics[width=\linewidth]{figures/trendestimation_details}

@@

\centerline{\bf Mechanistic forecasting}
\vspace*{2mm}
\begin{minipage}[t]{\linewidth}
    \includegraphics[width=\linewidth]{figures/forecasting_details}

@@

\centerline{\bf A common workflow to bridge trend estimation and forecasting} % jdavies15april: would it be possible to label this panel with components of the LPI to provide a clearer illustration of how the framework is implemented? % [FLAG: not sure I understood]
\vspace*{-3mm}
{\begin{multicols}{2}
\begin{minipage}[t]{\linewidth}
	\vfill
\includegraphics[width=\linewidth]{figures/iterativeworkflow_details}
\vfill
% \vspace{-0.5cm}\captionsetup{font=footnotesize}\captionof{figure}{Caption}\label{fig:trends}
\vspace*{3mm}
\end{minipage}

\columnbreak
\vspace*{1mm}
A universal workflow offers an opportunity to bridge statistical and process-based frameworks, integrating mechanistic knowledge and leveraging robust statistical approaches \citep[e.g.][]{rounce2020quantifying}. Process-based models would no longer be perceived as deterministic black boxes by other researchers but rather as robust statistical frameworks encapsulating both data structure and mechanistic knowledge. It would be an opportunity to spread the incorporation of mechanistic assumptions beyond the process-based modeling community, potentially improving trend estimates. For both trend estimation and forecasting, the workflow would refocus attention on the research question, highlighting the ecological hypotheses that justify the use and design of the model.

@@

\clearpage
\bibliography{forecastflows}

>>>>

\documentclass[11pt]{article}
\usepackage[top=0.70in, bottom=.80in, left=1in, right=1in]{geometry}
\renewcommand{\baselinestretch}{1.1}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{multicol}
\usepackage{tcolorbox}
\usepackage{caption}
\usepackage{todonotes}
\usepackage[font=sf]{caption}
\usepackage{hyperref}
\newcommand{\llabel}[1]{\hypertarget{lintarget:#1}{}\linelabel{lin:#1}}

@@

\noindent \emph{For:} Workflow for Applied Data Analysis theme issue for \emph{Phil Trans A} as an \emph{Opinion}
% They mention a tex template here: https://royalsocietypublishing.org/rsta/for-authors#question4
% Word length: I should check what they emailed but they say no more than 13 printed pages with 650 per page (whoa, we should be well below that! I am thinking 3-5K would be good)
% Already 5K!

@@

\begin{figure}[h]
	\centering
	\includegraphics{../figures/modeldataspaces_rotation}
	\caption{Different trajectories through model-data space during model development. Ideally, this begins in the pre-model/pre-data quadrant, where the research question is defined. The next step moves to the post-model/pre-data quadrant, where a model is build---independently of the data. This is where we should spend a good deal of time evaluating model behavior using simulated data. Once we are satisfied with our model, we can move to the post-model/post-data quadrant, where the model is fitted to real data. Then, we perform retrodictive checks to compare model predictions to observations, which likely give us some feedbacks to refine our model.}
	\label{fig:modeldata}
\end{figure}

@@

We argue a workflow that moves along the data-model space in a coherent sequence of steps (figure \ref{fig:modeldata}c) could reduce many of these problems and thus improve ecological science.
The first step of this workflow is to define an explicit research question and formulate hypotheses (step 1, figure \ref{fig:workflow}). This involves making clear assumptions about the most influential drivers, within the specific context of our study. This should guide the construction of a narrative model of how we believe the system works, focusing on the mechanisms that could generate the data we observe, including the observational error.  
\llabel{causalinf}This step of carefully formulating hypotheses has gained increasing attention in the ecological literature \citep{Grace2020}. From this narrative, we can then develop a mathematical model---an ensemble of equations that encapsulates our knowledge and is designed to answer our research question (step 2, figure \ref{fig:workflow}). The general idea is to start with a relatively simple model that we could refine later (see example worklow we provide). At this stage, prioritizing biologically meaningful parameters is crucial, as it allows us to have a sense of plausible parameter values. This means choosing a model formulation where each parameter corresponds to an interpretable behavior (which sometimes requires considering alternative parameterizations).

With a model in place, the next step focuses on testing and understanding it via data simulation (step 3, figure \ref{fig:workflow}). `Fake' or `test' data are generated directly from the model by fixing parameters to some reasonable values (which is straightforward if the parameters are interpretable) and from fake predictor data. 
We then fit our model to this simulated dataset and evaluate its ability to recover the prescribed values. At this stage, the focus is on understanding the model, so we may need to spend time thinking about whether each parameter is reasonable or not. This is also a step for making sure the model is working as expected (see example worklow). \llabel{datasimforall}Ideally, this data simulation step should be repeated several times with different parameter values, sampled within a reasonable range (or, in a Bayesian framework, from the prior distribution).

Once we are confident about our model structure, we can introduce real data as part of an initial model fitting step (step 4, figure \ref{fig:workflow}). This way, we obtain parameter estimates constrained by observations. 
These parameter estimates lead to the second data simulation step, this time using our fitted model parameters to generate predictions (step 5, figure \ref{fig:workflow}). This---which we call a retrodictive check (\citealp{betanworkflow}, also called a \llabel{PPC}posterior predictive check in a Bayesian workflow,  \citealp{Gelman2020})---allows model output to be compared to observations. 
It's only once all steps have been completed that we can interpret parameter values with respect to our research question.
The workflow encourages a focus on the full model, where any parameter (such as a trend estimate) must be carefully interpreted alongside others, as all are fundamental components that shape both inference and forecasting.

@@

\begin{figure}[h]
	\centering
	\vspace*{-0.1cm}
    \hspace*{-1.5cm}
	\includegraphics{../figures/figure_worflow_wsteps}
	\caption{The workflow we propose here \citep[which builds from recent advances in workflows,][]{betanworkflow,Gelman2020,Schad2020,grinsztajn2021,vandeschoot2021,Wolkovich2024} focuses on iterative feedbacks between the research question, model and data. Most research should start at Step 1, with the research question, followed by extensive model building and evaluation using simulated data (Steps 2 and 3) before proceeding to fitting the model on empirical data and examining it through retrodictive checks (Steps 4 and 5). Within this workflow, forecasting is a natural output of the process and not a separate process or one available for only certain modeling approaches.  The integration of observations (the data stream) occurs at Step 4---only after the model has been thoroughly evaluated---and can also highlight opportunities to collect new data and enrich the data stream.}
	\vspace*{-0.5cm}
	\label{fig:workflow}
\end{figure}

@@

\centerline{\bf Trend estimation}
\vspace*{2mm}
\begin{minipage}[t]{\linewidth}
    \includegraphics[width=\linewidth]{../figures/trendestimation_details}

@@

\centerline{\bf Mechanistic forecasting}
\vspace*{2mm}
\begin{minipage}[t]{\linewidth}
    \includegraphics[width=\linewidth]{../figures/forecasting_details}

@@

\centerline{\bf A common workflow to bridge trend estimation and forecasting} % jdavies15april: would it be possible to label this panel with components of the LPI to provide a clearer illustration of how the framework is implemented? % [FLAG: not sure I understood]
\vspace*{-3mm}
{\begin{multicols}{2}
\begin{minipage}[t]{\linewidth}
	\vfill
\includegraphics[width=\linewidth]{../figures/iterativeworkflow_details}
\vfill
% \vspace{-0.5cm}\captionsetup{font=footnotesize}\captionof{figure}{Caption}\label{fig:trends}
\vspace*{3mm}
\end{minipage}

\columnbreak
\vspace*{1mm}
A universal workflow offers an opportunity to bridge statistical and process-based frameworks, integrating mechanistic knowledge and leveraging robust statistical approaches \citep[e.g.][]{rounce2020quantifying}. Process-based models would no longer be perceived as deterministic black boxes by other researchers but rather as robust statistical frameworks encapsulating both data structure and mechanistic knowledge---and where the full model is fitted jointly. It would be an opportunity to spread the incorporation of mechanistic assumptions beyond the process-based modeling community, potentially improving trend estimates. For both trend estimation and forecasting, the workflow would refocus attention on the research question, highlighting the ecological hypotheses that justify the use and design of the model.

@@

\clearpage
\bibliography{../forecastflows.bib}
